Computational methods in landscape ecology

### 2024-10-22 Manuscript.qmd
Clean changes MJ
Move around some sentences
Comments Marti

format:
  pdf:
    template-partials:
      - title.tex
    include-in-header:
      text: |
        \usepackage{academicons}
        \usepackage{fontawesome5}
        \usepackage{lineno}\linenumbers
        \usepackage[noblocks]{authblk}
        \usepackage{setspace}
        \renewcommand\Affilfont{\footnotesize}
        \newcommand{\orcid}[1]{\href{https://orcid.org/#1}{\textcolor[HTML]{A6CE39}{\aiOrcid\hspace{1mm}#1}}}
        \newcommand{\mail}[1]{\href{mailto:#1}{\textcolor[HTML]{052E81}{\faEnvelope\hspace{1mm}#1}}}

@@

@@

{{< include Sections/01-introduction.qmd >}}
{{< include Sections/02-data.qmd >}}
{{< include Sections/03-spatial-patterns.qmd >}}
{{< include Sections/04-connectivity.qmd >}}
{{< include Sections/05-sampling.qmd >}}
{{< include Sections/06-simulations.qmd >}}
{{< include Sections/07-genetics.qmd >}}
{{< include Sections/08-spatial-planning.qmd >}}
{{< include Sections/09-conclusion.qmd >}}

@@

\footnotesize
\setstretch{1.0}
::: {#refs}
:::

>>>>

format:
  pdf:
    template-partials:
      - title.tex
    include-in-header:
      text: |
        \usepackage{academicons}
        \usepackage{fontawesome5}
        \usepackage{lineno}\linenumbers
        \usepackage[noblocks]{authblk}
        \usepackage{setspace}
        \renewcommand\Affilfont{\footnotesize}
        \newcommand{\orcid}[1]{\href{https://orcid.org/#1}{\textcolor[HTML]{A6CE39}{\aiOrcid\hspace{1mm}#1}}}
        \newcommand{\mail}[1]{\href{mailto:#1}{\textcolor[HTML]{32B2DA}{\faEnvelope\hspace{1mm}#1}}}

@@

\newpage

@@

{{< include Sections/01-introduction.qmd >}}
{{< include Sections/02-data.qmd >}}
{{< include Sections/03-spatial-patterns.qmd >}}
{{< include Sections/04-connectivity.qmd >}}
{{< include Sections/05-genetics.qmd >}}
{{< include Sections/06-sampling.qmd >}}
{{< include Sections/07-simulations.qmd >}}
{{< include Sections/08-spatial-planning.qmd >}}
{{< include Sections/09-conclusion.qmd >}}

@@

\footnotesize
::: {#refs}
:::

### 2024-10-22 Review.qmd
Move around some sentences
Comments Marti

\textcolor[HTML]{007895}{Reply: This is a very helpful comment about the general structure of the manuscript and we tried to address it the best possible. This included several points. First, we agree that a general section on data models is useful for readers. Thus, we added such a section at the beginning of the manuscript (L65 - 92). Second, we modified especially the first paragraphs of most (sub-)sections in the "Spatial patterns" section to improve the logical flow of the manuscript (L93 - 202). Third, we rearranged and modified the "Connectivity" section as well. We removed the "Structural connectivity" sub-section by including the key statements into the general paragraph about connectivity (L207 - 212).}

@@

\textcolor[HTML]{007895}{Reply: Indeed, we forgot to add a sub-section title which is added now (L116).}

@@

\textcolor[HTML]{007895}{Reply: We replaced "entropy" with "measure" (L128 - 130).}

@@

\textcolor[HTML]{007895}{Reply: We added a brief definition of edge length in this context (L144 - 146).}

@@

\textcolor[HTML]{007895}{Reply: We removed the term "query" from the sentence because we agree that it is a source of confusion. However, we believe that the sentence using only "areas" instead is clear enough for readers to follow (L193-201).}

@@

\textcolor[HTML]{007895}{Reply: We agree that the whole section was confusing for readers unfamiliar with the topic. Thus, we reworded it to make it easier to follow by removing a lot of the database jargon (L193-201).}

@@

\textcolor[HTML]{e93c26}{Reply: Thank you for pointing this out. We added it to the figure.}

@@

\textcolor[HTML]{007895}{Reply: The figure caption now includes more details how the different topcis of the figure are interlinked.}

@@

\textcolor[HTML]{007895}{Reply: We added "flow" as suggested (L204 - 205).}

@@

\textcolor[HTML]{007895}{Reply: We added "dispersal" as suggested (L217 - 219).}

@@

\textcolor[HTML]{007895}{Reply: We added dispersal as suggested (L217 - 219).}

@@

\textcolor[HTML]{007895}{Reply: We added the specification that movement also includes dispersal of plants. Furthermore, we believe that this aspect is now also highlighted due to the other suggestions of the reviewer that we implemented in this section. Generally, we added this each time movement is mentioned in the section (L204 - 284).}

@@

\textcolor[HTML]{007895}{Reply: We added "disperse" as suggested (L227 - 230).}

@@

\textcolor[HTML]{007895}{Reply: We are unsure why the point about edge artifacts was added to the manuscript. Thus, we completely removed it here since we think it adds more confusion than helping the readers (L230 - 232).}

@@

\textcolor[HTML]{007895}{Reply: We agree that the original statement about underlying assumptions was unclear and reworded the sentence (L237 - 239).}

@@

\textcolor[HTML]{007895}{Reply: We added this specification as suggested (L263 - 266).}

@@

\textcolor[HTML]{007895}{Reply: Thank you very much for adding this insightful remark about the references. We reworded the sentence to ensure the remarks are reflected in the manuscript (L331 - 334)}

@@

\textcolor[HTML]{007895}{Reply: We added that digital twins benefit from the increasing rate of data generation in ecology (L365 - 367).}

@@

\textcolor[HTML]{007895}{Reply: We renamed the section as suggested (L392).}

@@

\textcolor[HTML]{007895}{Reply: We reworded the sentence being more specific about the corresponding references (L408 - 410).}

@@

\textcolor[HTML]{007895}{Reply: Thank you for pointing out this detail. We added this as another potentail drawback of the method (L418 - 421).}

@@

\textcolor[HTML]{007895}{Reply: We reworded the sentence stating that there are differences between solvers and that the optimal solver depends on the specific problem. However, the best solver could potentially not be accessible due to software licenses (424 - 429).}

>>>>

\textcolor[HTML]{007895}{Reply: This is a very helpful comment about the general structure of the manuscript and we tried to address it the best possible. This included several points. First, we agree that a general section on data models is useful for readers. Thus, we added such a section at the beginning of the manuscript (L65 - 92). Second, we modified especially the first paragraphs of most (sub-)sections in the "Spatial patterns" section to improve the logical flow of the manuscript (L93 - 202). Third, we rearranged and modified the "Connectivity" section as well. We removed the "Structural connectivity" sub-section by including the key statements into the general paragraph about connectivity (L207 - 212). Last, we moved the "Landscape genetics" section closer to the "Connectivity" section due to the importance of connectivity modelling in landscape genetics.}

@@

\textcolor[HTML]{007895}{Reply: Indeed, we forgot to add a sub-section title which is added now (L115).}

@@

\textcolor[HTML]{007895}{Reply: We replaced "entropy" with "measure" (L127 - 129).}

@@

\textcolor[HTML]{007895}{Reply: We added a brief definition of edge length in this context (L143 - 145).}

@@

\textcolor[HTML]{007895}{Reply: We removed the term "query" from the sentence because we agree that it is a source of confusion. However, we believe that the sentence using only "areas" instead is clear enough for readers to follow (L195-197).}

@@

\textcolor[HTML]{007895}{Reply: We agree that the whole section was confusing for readers unfamiliar with the topic. Thus, we reworded it to make it easier to follow by removing a lot of the database jargon (L194-203).}

@@

\textcolor[HTML]{007895}{Reply: Thank you for pointing this out. We added it to the figure.}

@@

\textcolor[HTML]{007895}{Reply: The figure caption now includes more details how the different topics of the figure are interlinked.}

@@

\textcolor[HTML]{007895}{Reply: We added "flow" as suggested (L205 - 206).}

@@

\textcolor[HTML]{007895}{Reply: We added "dispersal" as suggested (L218 - 220).}

@@

\textcolor[HTML]{007895}{Reply: We added dispersal as suggested (L218 - 220).}

@@

\textcolor[HTML]{007895}{Reply: We added the specification that movement also includes dispersal of plants. Furthermore, we believe that this aspect is now also highlighted due to the other suggestions of the reviewer that we implemented in this section. Generally, we added this each time movement is mentioned in the section (L205 - 285).}

@@

\textcolor[HTML]{007895}{Reply: We added "disperse" as suggested (L228 - 231).}

@@

\textcolor[HTML]{007895}{Reply: We are unsure why the point about edge artifacts was added to the manuscript. Thus, we completely removed it here since we think it adds more confusion than helping the readers (L231 - 233).}

@@

\textcolor[HTML]{007895}{Reply: We agree that the original statement about underlying assumptions was unclear and reworded the sentence (L238 - 239).}

@@

\textcolor[HTML]{007895}{Reply: We added this specification as suggested (L264 - 267).}

@@

\textcolor[HTML]{007895}{Reply: Thank you very much for adding this insightful remark about the references. We reworded the sentence to ensure the remarks are reflected in the manuscript (L356 - 359)}

@@

\textcolor[HTML]{007895}{Reply: We added that digital twins benefit from the increasing rate of data generation in ecology (L384 - 386).}

@@

\textcolor[HTML]{007895}{Reply: We renamed the section as suggested (L395).}

@@

\textcolor[HTML]{007895}{Reply: We reworded the sentence being more specific about the corresponding references (L411 - 413).}

@@

\textcolor[HTML]{007895}{Reply: Thank you for pointing out this detail. We added this as another potential drawback of the method (L421 - 424).}

@@

\textcolor[HTML]{007895}{Reply: We reworded the sentence stating that there are differences between solvers and that the optimal solver depends on the specific problem. However, the best solver could potentially not be accessible due to software licenses (427 - 432).}

### 2024-10-22 Sections/03-spatial-patterns.qmd
Minor typos
Adding line numbers to replies
Move around some sentences
Comments Marti

Traditionally and at present, many prominent approaches that quantify spatial patterns revolve around raster data using categorical values based on the patch mosaic model [@Frazier2017LandscapeMetrics; @Costanza2019DescribingAnalyzing].
<!-- R1: They are also naturally extracted from remote sensing products (raster data), which often makes them easier to monitor over time -->
The strength of landscape metrics is that they are easy to apply, communicate, and can be calculated straightforwardly from raster data based on remote sensing derived products [@Dramstad2009SpatialMetrics; @Lausch2015UnderstandingQuantifying].
However, also limitations have been identified for many landscape metrics.
These limitations include shortcomings to quantify the spatial structure, sensitivity to both spatial scale and thematic resolution, and correlation and redundancy between metrics [@Li2004UseMisuse; @Dramstad2009SpatialMetrics; @Kupfer2012LandscapeEcology].

To address limitations related to correlation and redundancy several approaches have been used to identify core metrics that capture main components of landscape patterns, such as multivariate factor analysis [@Riitters1995FactorAnalysis], multivariate statistics [@Lustig2015MorePredictable], principal component analysis [@Cushman2008ParsimonyLandscape], or the variance inflation factor [@Plexida2014SelectingLandscape].
More recently, principal component analysis over a set of landscape blocks revealed two main components of landscape configuration, namely complexity and aggregation which together explain about 70% of variance [@Nowosad2018GlobalInventory].
These components of landscape patterns are in line with recent reviews that have similarly suggested that there are two fundamental components of landscape patterns, namely amount and adjacency [@Gustafson2019HowHas; @Riitters2019PatternMetrics].
These components are connected to complexity and aggregation as suggested by @Nowosad2018GlobalInventory and later formalized in @Nowosad2019InformationTheory.

@@

Surface metrics are based on the gradient surface model using raster data and continuous values [@McGarigal2009SurfaceMetrics; @Cushman2010GradientParadigm]. 
These metrics are mostly adapted from microscopy and molecular physics [@McGarigal2009SurfaceMetrics; @Kedron2018SurfaceMetrics].
The gradient surface model can increase the resemblance of the data to the natural world because it allows for the inclusion of more heterogeneity within each grid cell [@Cushman2010GradientParadigm; @Kedron2018SurfaceMetrics].
Many surface metrics have analogous landscape metrics [@Cushman2010GradientParadigm], however, using surface metrics allows for the exploration of different or additional patterns and potentially pattern-process links [@McGarigal2009SurfaceMetrics; @Borthwick2020AlternativeQuantifications].
The metrics are able to quantify various characteristics, such as roughness, skewness and kurtosis, total and relative amplitudes, curvatures of local peaks, or surface bearing area ratios [@Cushman2010GradientParadigm].
<!-- R1: I guess that such metrics are also very scale dependent -->
Nevertheless, similar to landscape metrics, also surface metrics are scale-dependent because they are used to quantify landscape heterogeneity which is itself scale-dependent [@Frazier2016SurfaceMetrics].
Software to calculate these metrics is still rare and further research is needed into the specific pattern-process links and ecologically meaningful interpretations [@Kedron2018SurfaceMetrics; @Frazier2019EmergingTrajectories].
Related to surface metrics, another recent approach to quantifying the gradient surface model is based on frequencies and local adjacencies of continuous input pixel values [@Riitters2023MappingLandscape].

## Vector metrics

@@

As hundreds of metrics have already been developed for gridded raster data, the most straightforward approach may be to reimplement these same metrics for vector data. 
Although this is possible for many metrics, there are also metrics specifically related to the corresponding data models [@Yao2022VecLIFramework].
For example, in urban planning vector-based metrics are applied to quantify the shapes of urban areas and characterize the complexity of building footprints [@Basaraner2017PerformanceShape].
Earlier approaches to quantifying shape complexity included four categories of compactness measures: perimeter-area, single parameters of related circles, dispersion of elements of the area around a centroid, and direct comparison to standard shapes [@Maceachren1985CompactnessGeographic].
More recently, a unified theoretical foundation for measuring shape compactness was introduced using a set of ten distinct properties of a circle and metrics associated with each of these properties [@Angel2010TenCompactness].
However, also vector-based metrics have limitations.
The most important limitation relates to computational complexity, which makes calculations of vector-based metrics slower than their raster equivalents.
Another technical issue is the requirement for topologically correct data (e.g., geometries cannot overlap), which is often problematic, especially for data from different sources.
Last, the pattern-process link for vector-based metrics remains underexplored compared to raster-based metrics.

@@

Spatial patterns can be analyzed through a range of computational operations, e.g., comparing, searching, or grouping.
These operations are based on spatial signatures (multi-numerical representations of landscape pattern) and dissimilarity measures (functions that quantify differences between the signatures), and can be calculated for different areas or for the same area at two different moments in time.
Comparing spatial signatures is often used to analyze landscape dynamics, e.g., to detect changes in landscape structure over time [@Duncan2023MonitoringSpatial].
<!-- R1: As defined here, it looks to me that this operation is not necessarily linked to a query area, but rather to a spatial signature which can be directly specified or extracted from a query area -->
Furthermore, signature-based searches can be used to compare the spatial signature of a focal area to the spatial signatures of multiple other areas.
This allows for the identification of areas with similar signatures compared to the focal area, e.g., areas with similar environmental conditions [@Jasiewicz2014LandscapeSimilarity].

Additionally, it is possible to calculate spatial signatures for multiple areas and group them to similar clusters based on their signatures [@Nowosad2021PatternbasedIdentification].

>>>>

Traditionally and at present, many prominent approaches that quantify spatial patterns revolve around raster data using categorical values based on the patch mosaic model [@Frazier2017LandscapeMetrics; @Costanza2019DescribingAnalyzing].
<!-- R1: They are also naturally extracted from remote sensing products (raster data), which often makes them easier to monitor over time -->
The strength of landscape metrics is that they are easy to apply, communicate, and can be calculated straightforwardly from raster data based on remote sensing derived products [@Dramstad2009SpatialMetrics; @Lausch2015UnderstandingQuantifying].
However, limitations have also been identified for many landscape metrics.
These include shortcomings to quantify the spatial structure, sensitivity to both spatial scale and thematic resolution, and correlation and redundancy between metrics [@Li2004UseMisuse; @Dramstad2009SpatialMetrics; @Kupfer2012LandscapeEcology].

To address limitations related to correlation and redundancy several approaches have been used to identify core metrics that capture main components of landscape patterns, such as multivariate factor analysis [@Riitters1995FactorAnalysis], multivariate statistics [@Lustig2015MorePredictable], principal component analysis [@Cushman2008ParsimonyLandscape], or the variance inflation factor [@Plexida2014SelectingLandscape].
More recently, principal component analysis over a set of landscape blocks revealed two main components of landscape configuration, namely complexity and aggregation which together explain about 70% of variance [@Nowosad2018GlobalInventory].
These results are in line with recent reviews that have similarly suggested that there are two fundamental components of landscape patterns, namely amount and adjacency [@Gustafson2019HowHas; @Riitters2019PatternMetrics].
These components are connected to complexity and aggregation as suggested by @Nowosad2018GlobalInventory and later formalized in @Nowosad2019InformationTheory.

@@

Surface metrics are based on the gradient surface model using raster data and continuous values [@McGarigal2009SurfaceMetrics; @Cushman2010GradientParadigm]. 
These metrics are mostly adapted from microscopy and molecular physics [@McGarigal2009SurfaceMetrics; @Kedron2018SurfaceMetrics].
The gradient surface model can increase the resemblance of the data to the natural world because it allows for the inclusion of more heterogeneity within each grid cell [@Cushman2010GradientParadigm; @Kedron2018SurfaceMetrics].
Many surface metrics have analogous landscape metrics [@Cushman2010GradientParadigm], however, using surface metrics allows for the exploration of different or additional patterns and potentially pattern-process links [@McGarigal2009SurfaceMetrics; @Borthwick2020AlternativeQuantifications].
The metrics are able to quantify various characteristics, such as roughness, skewness and kurtosis, total and relative amplitudes, curvatures of local peaks, or surface bearing area ratios [@Cushman2010GradientParadigm].

<!-- R1: I guess that such metrics are also very scale dependent -->
Nevertheless, similar to landscape metrics, also surface metrics are scale-dependent because they are used to quantify landscape heterogeneity which is itself scale-dependent [@Frazier2016SurfaceMetrics].
Software to calculate these metrics is still rare and further research is needed into the specific pattern-process links and ecologically meaningful interpretations [@Kedron2018SurfaceMetrics; @Frazier2019EmergingTrajectories].

Related to surface metrics, another recent approach to quantifying the gradient surface model is based on frequencies and local adjacencies of continuous input pixel values [@Riitters2023MappingLandscape].

## Vector-based metrics

@@

As hundreds of metrics have already been developed for gridded raster data, the most straightforward approach may be to reimplement these same metrics for vector data. 
Although this is possible for many metrics, there are also metrics specifically related to the corresponding data models [@Yao2022VecLIFramework].
For example, in urban planning vector-based metrics are applied to quantify the shapes of urban areas and characterize the complexity of building footprints [@Basaraner2017PerformanceShape].
Earlier approaches to quantifying shape complexity included four categories of compactness measures: perimeter-area, single parameters of related circles, dispersion of elements of the area around a centroid, and direct comparison to standard shapes [@Maceachren1985CompactnessGeographic].
More recently, a unified theoretical foundation for measuring shape compactness was introduced using a set of ten distinct properties of a circle and metrics associated with each of these properties [@Angel2010TenCompactness].

However, also vector-based metrics have limitations.
The most important limitation relates to computational complexity, which makes calculations of vector-based metrics slower than their raster equivalents.
Another technical issue is the requirement for topologically correct data (e.g., geometries cannot overlap), which is often problematic, especially for data from different sources.
Last, the pattern-process link for vector-based metrics remains underexplored compared to raster-based metrics.

@@

Spatial patterns can be analyzed through a range of computational operations, e.g., comparing, searching, or grouping.
These operations are based on spatial signatures (multi-numerical representations of landscape pattern) and dissimilarity measures (functions that quantify differences between the signatures), and can be calculated for different areas or for the same area at two different moments in time.
Comparing spatial signatures is often used to analyze landscape dynamics, e.g., to detect changes in landscape structure over time [@Duncan2023MonitoringSpatial].
<!-- R1: As defined here, it looks to me that this operation is not necessarily linked to a query area, but rather to a spatial signature which can be directly specified or extracted from a query area -->
Furthermore, signature-based searches can be used to compare the spatial signature of a focal area to the spatial signatures of multiple other areas.
This allows for the identification of areas with similar signatures compared to the focal area, e.g., areas with similar environmental conditions [@Jasiewicz2014LandscapeSimilarity].
Additionally, it is possible to calculate spatial signatures for multiple areas and group them to similar clusters based on their signatures [@Nowosad2021PatternbasedIdentification].

### 2024-10-22 Sections/09-conclusion.qmd
Move around some sentences
Comments Marti

<!-- MH: Double check programming languages -->

| Topic | Programming language | Software | Reference |
|:---:|:---:|:---:|:---:|
| Spatial patterns | R | sf | [@Pebesma2018SimpleFeatures] |
| Spatial patterns | R | terra | [@Hijmans2024TerraSpatial] |
| Spatial patterns | R | landscapemetrics | [@Hesselbarth2019LandscapemetricsOpensource] |
| Spatial patterns | R | multilandr | [@Huais2024MultilandrPackage] |
| Spatial patterns | R | bespatial | [@Nowosad2024BespatialBoltzmann] |
| Spatial patterns | R | rasterdiv | [@Rocchini2021RasterdivInformation] |
| Spatial patterns | R | LandComp | [@Konrad2023AdaptingMultiscale] |
| Spatial patterns | R | geodiv | [@Smith2021GeodivPackage] |
| Spatial patterns | R | glcm | [@Zvoleff2020GlcmCalculate] |
| Spatial patterns | R | vectormetrics | [@Matuszek2024VectormetricsLandscape] |
| Spatial patterns | R | motif | [@Nowosad2021MotifOpensource] |
| Spatial patterns | Python | GeoPandas | [@Jordahl2020GeopandasPython] |
| Spatial patterns | Python | Rasterio | [@Gillies2013RasterioGeospatial] |
| Spatial patterns | Python | PyLandStats | [@Bosch2019PyLandStatsOpensource] |
| Spatial patterns | Python | LecoS | [@Jung2016LecoSPython] |
| Spatial patterns | Julia | GeoStats | [@Hoffimann2018GeoStatsjlHighperformance] |
| Spatial patterns | Julia | GeoInterface | [@JuliaGeoandcontributors2024GeoInterfaceJulia] |
| Spatial patterns | Stand-alone software | FRAGSTATS | [@McGarigal2023FRAGSTATSV4] |
| Spatial patterns | Stand-alone software | GuidosToolbox | [@Vogt2017GuidosToolboxUniversal] |
| Connectivity | R | ResistanceGA | [@Peterman2018ResistanceGAPackage] |
| Connectivity | R | lconnect | [@Mestre2023LconnectPackage] |
| Connectivity | R | grainscape | [@Chubaty2020ToolboxGrainscape] |
| Connectivity | Julia | Circuitscape | [@Anantharaman2019CircuitscapeJulia] |
| Connectivity | Julia | ConScape | [@vanMoorter2023AcceleratingAdvances] |
| Connectivity | Julia | Omniscape | [@Landau2021OmniscapejlSoftware] |
| Connectivity | Stand-alone software | Graphab | [@Foltete2012SoftwareTool] |
| Connectivity | Stand-alone software | Conefor | [@Saura2009ConeforSensinode] |
| Connectivity | Stand-alone software | LSCorridors | [@Ribeiro2017LandScapeCorridors] |
| Sampling | Python | ZonalMetrics | [@Adamczyk2017ZonalMetricsPython] |
| Simulations | R | NLMR | [@Sciaini2018NLMRLandscapetools] |
| Simulations | R | rflsgen | [@Justeau-Allaire2022FragmentedLandscape] |
| Simulations | Python | NLMpy | [@Etherington2015NLMpyPython] |
| Simulations | Python | Pathwalker | [@UnnithanKumar2022PathwalkerNew] |
| Simulations | Julia | NeutralLandscapes | [@Poisot2023NeutralLandscapesJulia] |
| Simulations | NetLogo | EFForTS-LGraf | [@Salecker2019EFForTSLGrafLandscape] |
| Simulations | Stand-alone software | GradientLand | [@Cambui2015GradientLandSoftware] |
| Simulations | Stand-alone software | Landscape Generator | [@vanStrien2016ImprovedNeutral] |
| Simulations | Stand-alone software | HexSim | [@Schumaker2018HexSimModeling] |
| Simulations | Stand-alone software | RangeShifter2.0 | [@Bocedi2021RangeShifter20] |
| Genetics | R | graph4lg | [@Savary2021Graph4lgPackage] |
| Genetics | R | landgenreport | [@Gruber2015LandgenreportNew] |
| Genetics | R | adegenet | [@Jombart2008AdegenetPackage] |
| Genetics | R | ape | [@Paradis2019Ape50] |
| Genetics | Python | Python based scripts | [@Etherington2011PythonBased] |
| Genetics | Python | SDMtoolbox 2.0 | [@Brown2017SDMtoolbox20] |
| Planning | R | prioritizr | [@Hanson2024PrioritizrSystematic] |
| Planning | R | CoCo | [@vanMantgem2023CocoConservation] |
| Planning | R | restoptr | [@Justeau-Allaire2023RestoptrPackage] |
| Planning | Python | CAPTAIN | [@Silvestro2022ImprovingBiodiversity] |

@@

>>>>

\newpage
\setstretch{1.0}

| Topic | Programming language | Software | Reference |
|:---:|:---:|:---:|:---:|
| Spatial patterns | R | sf | [@Pebesma2018SimpleFeatures] |
| Spatial patterns | R | terra | [@Hijmans2024TerraSpatial] |
| Spatial patterns | R | landscapemetrics | [@Hesselbarth2019LandscapemetricsOpensource] |
| Spatial patterns | R | multilandr | [@Huais2024MultilandrPackage] |
| Spatial patterns | R | bespatial | [@Nowosad2024BespatialBoltzmann] |
| Spatial patterns | R | rasterdiv | [@Rocchini2021RasterdivInformation] |
| Spatial patterns | R | LandComp | [@Konrad2023AdaptingMultiscale] |
| Spatial patterns | R | geodiv | [@Smith2021GeodivPackage] |
| Spatial patterns | R | glcm | [@Zvoleff2020GlcmCalculate] |
| Spatial patterns | R | vectormetrics | [@Matuszek2024VectormetricsLandscape] |
| Spatial patterns | R | motif | [@Nowosad2021MotifOpensource] |
| Spatial patterns | Python | GeoPandas | [@Jordahl2020GeopandasPython] |
| Spatial patterns | Python | Rasterio | [@Gillies2013RasterioGeospatial] |
| Spatial patterns | Python | PyLandStats | [@Bosch2019PyLandStatsOpensource] |
| Spatial patterns | Python | LecoS | [@Jung2016LecoSPython] |
| Spatial patterns | Julia | GeoStats | [@Hoffimann2018GeoStatsjlHighperformance] |
| Spatial patterns | Julia | GeoInterface | [@JuliaGeoandcontributors2024GeoInterfaceJulia] |
| Spatial patterns | Stand-alone software | FRAGSTATS | [@McGarigal2023FRAGSTATSV4] |
| Spatial patterns | Stand-alone software | GuidosToolbox | [@Vogt2017GuidosToolboxUniversal] |
| Connectivity | R | ResistanceGA | [@Peterman2018ResistanceGAPackage] |
| Connectivity | R | lconnect | [@Mestre2023LconnectPackage] |
| Connectivity | R | grainscape | [@Chubaty2020ToolboxGrainscape] |
| Connectivity | Julia | Circuitscape | [@Anantharaman2019CircuitscapeJulia] |
| Connectivity | Julia | ConScape | [@vanMoorter2023AcceleratingAdvances] |
| Connectivity | Julia | Omniscape | [@Landau2021OmniscapejlSoftware] |
| Connectivity | Stand-alone software | Graphab | [@Foltete2012SoftwareTool] |
| Connectivity | Stand-alone software | Conefor | [@Saura2009ConeforSensinode] |
| Connectivity | Stand-alone software | LSCorridors | [@Ribeiro2017LandScapeCorridors] |
| Landscape genetics | R | graph4lg | [@Savary2021Graph4lgPackage] |
| Landscape genetics | R | landgenreport | [@Gruber2015LandgenreportNew] |
| Landscape genetics | R | adegenet | [@Jombart2008AdegenetPackage] |
| Landscape genetics | R | ape | [@Paradis2019Ape50] |
| Landscape genetics | Python | Python based scripts | [@Etherington2011PythonBased] |
| Landscape genetics | Python | SDMtoolbox 2.0 | [@Brown2017SDMtoolbox20] |
| Sampling | Python | ZonalMetrics | [@Adamczyk2017ZonalMetricsPython] |
| Simulation models | R | NLMR | [@Sciaini2018NLMRLandscapetools] |
| Simulation models | R | rflsgen | [@Justeau-Allaire2022FragmentedLandscape] |
| Simulation models | Python | NLMpy | [@Etherington2015NLMpyPython] |
| Simulation models | Python | Pathwalker | [@UnnithanKumar2022PathwalkerNew] |
| Simulation models | Julia | NeutralLandscapes | [@Poisot2023NeutralLandscapesJulia] |
| Simulation models | Stand-alone software | GradientLand | [@Cambui2015GradientLandSoftware] |
| Simulation models | Stand-alone software | Landscape Generator | [@vanStrien2016ImprovedNeutral] |
| Simulation models | Stand-alone software | HexSim | [@Schumaker2018HexSimModeling] |
| Simulation models | Stand-alone software | RangeShifter2.0 | [@Bocedi2021RangeShifter20] |
| Spatial planning | R | prioritizr | [@Hanson2024PrioritizrSystematic] |
| Spatial planning | R | CoCo | [@vanMantgem2023CocoConservation] |
| Spatial planning | R | restoptr | [@Justeau-Allaire2023RestoptrPackage] |
| Spatial planning | Python | CAPTAIN | [@Silvestro2022ImprovingBiodiversity] |

@@

\newpage
\normalsize
\setstretch{1.5}
